\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}

\setlength{\parskip}{0.8em}
\setlength{\parindent}{0pt}

\begin{document}

\noindent Dear Professor Olwyn Byron,
\\
\\
Thank you for handling our manuscript ``From Rule-Taking to Rule-Making Chemistry: A Minimal Population-Environment Feedback Model'' and for the opportunity to revise it. We are grateful that both reviewers found the work relevant and recommended publication.

We have carefully addressed all comments raised by both reviewers. The most substantial changes concern the chemical grounding and scope of the Discussion. Following the reviewers' suggestions, we have significantly expanded the literature on replicator--environment feedback, incorporating both experimental work on synthetic replicator ecosystems and theoretical frameworks including game-theoretic models with environmental feedbacks and metabolically coupled replicator systems. We have also added a new Discussion paragraph on molecular complexity, engaging with the references recommended by Reviewer~2 and framing the connection between complexity measures and our feedback threshold as a compelling direction for future work. A further addition discusses the role of classical reaction mechanism analysis in constraining the model's abstract parameters.

On the modelling side, we corrected the modifier-production reaction to explicitly include the substrate and added concrete chemical examples of processes this reaction represents. We also clarified the purpose of the quasi-steady environment limit, emphasising its role as a baseline that reveals the qualitatively new behaviour introduced by environmental persistence.

Several smaller clarifications were made throughout the Model section in response to Reviewer~1, including the meaning of ``frequency'', the origin-of-life rationale for the chemostat assumption, the role of relative fitness differences in the replicator equation, and the early introduction of the parameter~$p$.

A point-by-point response to both reviewers is enclosed separately. 

We hope that the revisions have satisfactorily addressed the reviewers' concerns and that the manuscript is now suitable for publication in \textit{Royal Society Interface}.

\noindent Sincerely,
\\
\\
\\
\\
\\
\\
\\
\noindent Celia Blanco\\
\small Centro de Astrobiolog\'ia (CSIC-INTA)\\
\small \texttt{celia.blanco@cab.inta-csic.es}

\end{document}
